---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd. -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
run_converter <- identical(Sys.getenv("OPENDATALOADER_README_RUN"), "true")
```

# opendataloader

<!-- badges: start -->
[![R-CMD-check](https://github.com/muschellij2/opendataloader/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/muschellij2/opendataloader/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/muschellij2/opendataloader/graph/badge.svg)](https://app.codecov.io/gh/muschellij2/opendataloader)
<!-- badges: end -->

`opendataloader` is an R interface to the Python
[`opendataloader-pdf`](https://github.com/opendataloader-project/opendataloader-pdf)
package. It converts PDFs to Markdown, JSON, and HTML through
[`reticulate`](https://rstudio.github.io/reticulate/).

The package uses `reticulate::py_require()` so recent versions of reticulate
can resolve `opendataloader-pdf` in a managed Python environment. The upstream
Python package requires Python 3.10 or newer and Java 11 or newer.

## Installation

```{r install, eval = FALSE}
install.packages("remotes")
remotes::install_github("muschellij2/opendataloader")
```

If you want reticulate to use a managed Python environment for this session,
set this before loading the package:

```{r managed-python, eval = FALSE}
Sys.setenv(RETICULATE_PYTHON = "managed")
library(opendataloader)
```

## Basic Use

The package includes a small example PDF so examples can use `system.file()`
instead of relying on local files.

```{r example-pdf}
library(opendataloader)

example_pdf <- system.file(
  "extdata",
  "opendataloader-example.pdf",
  package = "opendataloader"
)

if (!nzchar(example_pdf)) {
  example_pdf <- file.path("inst", "extdata", "opendataloader-example.pdf")
}

file.exists(example_pdf)
```

Convert one PDF to Markdown and JSON:

```{r convert-one, eval = run_converter}
output_dir <- tempfile("opendataloader-output-")
dir.create(output_dir)

conversion <- convert_pdf(
  input_path = example_pdf,
  output_dir = output_dir,
  format = c("markdown", "json"),
  verbose = TRUE
)

conversion
conversion$output_files
conversion$result
```

Set `verbose = FALSE` to ask the Python converter to run quietly:

```{r quiet-convert, eval = FALSE}
quiet_conversion <- convert_pdf(
  input_path = example_pdf,
  output_dir = output_dir,
  format = "markdown",
  verbose = FALSE
)
```

## Conversion Options

Common Python converter options are exposed as typed R arguments so invalid
values fail before reticulate calls Python:

```{r conversion-options, eval = FALSE}
conversion <- convert_pdf(
  input_path = example_pdf,
  output_dir = output_dir,
  format = c("markdown", "html"),
  pages = "1-3",
  sanitize = TRUE,
  keep_line_breaks = TRUE,
  reading_order = "xycut",
  table_method = "cluster",
  image_output = "external",
  image_format = "png",
  include_header_footer = FALSE,
  detect_strikethrough = TRUE,
  verbose = FALSE
)
```

Encrypted PDFs can be converted with `password`:

```{r encrypted-pdf, eval = FALSE}
convert_pdf(
  input_path = "encrypted.pdf",
  output_dir = "converted",
  password = "secret",
  format = "markdown"
)
```

Advanced or newly-added Python options can still be passed through `...`, but
they must be named and must be simple R values (`NULL`, logical, numeric, or
character).

By default, README rendering does not run the converter because that requires
Java and the Python dependency resolver. To run the conversion chunks while
rendering locally:

```{r render-readme, eval = FALSE}
Sys.setenv(OPENDATALOADER_README_RUN = "true")
devtools::build_readme()
```

## Batch Conversion

Batch conversion is usually faster than converting files one at a time because
the upstream Python package can handle multiple files in a single call.

```{r batch-convert, eval = run_converter}
pdfs <- rep(example_pdf, 2)
batch_dir <- tempfile("opendataloader-batch-")
dir.create(batch_dir)

batch_conversion <- convert_pdf(
  input_path = pdfs,
  output_dir = batch_dir,
  format = c("markdown", "json")
)

batch_conversion$output_files
```

For your own documents, pass a vector of file paths:

```{r own-files, eval = FALSE}
pdfs <- list.files("pdfs", pattern = "[.]pdf$", full.names = TRUE)

conversion <- convert_pdf(
  input_path = pdfs,
  output_dir = "converted",
  format = c("markdown", "json", "html")
)

conversion$output_files
conversion$result
```

## Optional Hybrid Mode

Hybrid mode is an optional Python extra. Declare it before Python is initialized
in the R session:

```{r hybrid-require, eval = FALSE}
opendataloader_pdf_require(extras = "hybrid")
```

Then pass the hybrid backend through to the converter:

```{r hybrid-convert, eval = FALSE}
convert_pdf(
  input_path = example_pdf,
  output_dir = "converted-hybrid",
  hybrid = "docling-fast",
  hybrid_mode = "full"
)
```

## Python Requirements

Inspect the Python requirements reticulate has collected:

```{r requirements}
opendataloader_python_requirements()
```

Use `py_require_opendataloader_pdf()` as an alias if you prefer names that make
the reticulate connection explicit:

```{r py-require-alias, eval = FALSE}
py_require_opendataloader_pdf(
  extras = "hybrid",
  python_version = ">=3.10"
)
```

## Java Check

`check_java()` verifies that `java` is available and reports Java 11 or newer:

```{r java-check, eval = FALSE}
check_java()
```

If this fails, install Java 11 or newer and make sure `java` is on `PATH`.

## Troubleshooting

If reticulate selects a Python environment that does not have
`opendataloader-pdf`, start a fresh R session and request a managed Python
environment before loading this package:

```{r troubleshooting-managed, eval = FALSE}
Sys.setenv(RETICULATE_PYTHON = "managed")
library(opendataloader)
```

If dependency resolution is unavailable because you are offline, configure
reticulate or `uv` before Python is initialized:

```{r troubleshooting-uv, eval = FALSE}
Sys.setenv(UV_OFFLINE = "1")
library(opendataloader)
```

For reproducible examples or tests, use the bundled example PDF:

```{r system-file-summary}
basename(example_pdf)
```
