---
title: "Using opendataloader"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using opendataloader}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`opendataloader` is a small R wrapper around the Python
`opendataloader-pdf` package. The Python package is declared through
`reticulate::py_require()` when this package loads, so recent versions of
reticulate can resolve a managed Python environment automatically.

The upstream project recommends batch conversion because each conversion starts
a Java process. The same pattern in R is to pass a vector of PDF paths in one
call:

```{r batch-convert, eval = FALSE}
library(opendataloader)

pdf_files <- c(
  "samples/pdf/1901.03003.pdf",
  "samples/pdf/2408.02509v1.pdf",
  "samples/pdf/chinese_scan.pdf",
  "samples/pdf/lorem.pdf"
)

output_dir <- tempfile("opendataloader-output-")
dir.create(output_dir)

convert_pdf(
  input_path = pdf_files,
  output_dir = output_dir,
  format = c("json", "markdown"),
  quiet = TRUE
)

list.files(output_dir)
```

Example output from the upstream batch-processing example:

```text
Found 4 PDFs in pdf/

==========================================================
Method 1: Batch convert with file list
==========================================================
Document                                  Pages Top-level
----------------------------------------------------------
1901.03003                                   15       241
2408.02509v1                                 14       365
chinese_scan                                  1         1
lorem                                         1         2
----------------------------------------------------------
Total                                        31       609
Processed 4 documents
Time: 7.95s (single JVM invocation)
```

Hybrid mode uses an optional Python extra and should be requested before Python
is initialized:

```{r hybrid, eval = FALSE}
opendataloader_pdf_require(extras = "hybrid")

convert_pdf(
  input_path = "samples/pdf/1901.03003.pdf",
  output_dir = output_dir,
  hybrid = "docling-fast",
  hybrid_mode = "full"
)
```

The Python package requires Python 3.10 or newer and Java 11 or newer.
